---
title: "Seed bank"
subtitle: "Seeds and cuttings from the collection"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{=html}
<style>#quarto-header,.navbar{display:none !important;}</style>
<div class="en-nav">
  <span class="brand">Arboretum</span>
  <a href="index.html">Home</a>
  <a href="collection.html" class="current">The Collection</a>
  <a href="map.html">Map</a>
  <a href="self-guided-tour.html">Self-guided tour</a>
  <a href="wellbeing.html">Wellbeing</a>
  <a href="visit.html">Visit</a>
  <a href="sponsor.html">Sponsor a tree</a>
  <a href="donate.html">Donate</a>
  <a href="companies.html">Companies</a>
  <a href="contact.html">Contact</a>
  <span class="spacer"></span>
  <a href="experiences.html" class="cta">Book a visit</a>
  <a href="../banco_semillas.html" class="lang">Español ▸</a>
</div>
```

```{r}
#| label: setup
source("scripts/cargar_datos.R")   # arboles + helpers

esc <- function(x){x<-as.character(x);x<-gsub("&","&amp;",x,fixed=TRUE);x<-gsub("<","&lt;",x,fixed=TRUE);x<-gsub(">","&gt;",x,fixed=TRUE);gsub('"',"&quot;",x,fixed=TRUE)}
opciones <- function(v) paste0("<option value=\"\">All</option>", paste0("<option value=\"",esc(v),"\">",esc(v),"</option>",collapse=""))

bs <- read_excel("data/banco_semillas.xlsx", sheet = "Semillas") %>%
  mutate(
    familia           = ifelse(tiene_valor(familia), familia, "—"),
    nombre_comun      = ifelse(tiene_valor(nombre_comun), nombre_comun, ""),
    nombre_cientifico = ifelse(tiene_valor(nombre_cientifico), nombre_cientifico, ""),
    bkey              = binomio(nombre_cientifico),
    iucn              = iucn_de(nombre_cientifico)
  )
bs$con_html <- iucn_badge(bs$iucn, "en")

col_ori <- if ("continente_norm" %in% names(arboles)) "continente_norm" else "continente"
lut <- arboles %>%
  mutate(bkey = binomio(nombre_cientifico)) %>%
  filter(nzchar(bkey)) %>%
  transmute(bkey, slug, ori = ifelse(tiene_valor(.data[[col_ori]]), .data[[col_ori]], "")) %>%
  distinct(bkey, .keep_all = TRUE)
bs <- bs %>% left_join(lut, by = "bkey")
bs$ori[is.na(bs$ori)] <- ""
bs$slug[is.na(bs$slug)] <- ""

n_acc <- nrow(bs)
n_fam <- length(unique(bs$familia[bs$familia != "—"]))
n_esp <- length(unique(bs$bkey[nzchar(bs$bkey)]))

try({
  dir.create("recursos", showWarnings = FALSE)
  out <- bs %>% transmute(id, familia, nombre_cientifico, nombre_comun, origin = ori, conservation = iucn)
  write.csv(out, "recursos/banco_semillas.csv", row.names = FALSE, fileEncoding = "UTF-8")
}, silent = TRUE)

sel_con <- paste0("<option value=\"\">All</option>",
  paste0(vapply(c("EN","VU","NT","LC"), function(k) sprintf("<option value=\"%s\">%s</option>", k, iucn_label(k,"en")), character(1)), collapse=""))
```

The arboretum conserves **`r n_acc` accessions** of seeds and cuttings — a basis for
**propagation**, **restoration**, **education** and **exchange** with other botanical
institutions. Each species links to its profile; where available we show its
**conservation status (IUCN)** and its **origin**. Search by name or combine the filters.

<p style="margin:.6rem 0 0;">
  <a href="../recursos/banco_semillas.csv" download class="btn btn-outline-success btn-sm" role="button"><i class="bi bi-download"></i> Download catalogue (CSV)</a>
  <span style="font-size:.8rem;color:#777;margin-left:.4rem;">Open data · free to use with attribution</span>
</p>

```{r results='asis'}
cat(sprintf('
<style>
  .bs-stats{display:flex;flex-wrap:wrap;gap:.8rem;margin:1.1rem 0;}
  .bs-stat{background:#f4f8f5;border-radius:10px;padding:.55rem 1.1rem;text-align:center;min-width:110px;}
  .bs-stat .n{display:block;font-size:1.6rem;font-weight:700;color:#B8860B;line-height:1.1;}
  .bs-stat .l{font-size:.78rem;color:#555;}
  .bs-controles{display:flex;flex-wrap:wrap;gap:.6rem;align-items:end;margin:1rem 0;}
  .bs-controles label{font-size:.8rem;color:#555;display:block;}
  .bs-controles input,.bs-controles select{padding:.4rem .5rem;border:1px solid #ccc;border-radius:6px;}
  #bs-q{min-width:220px;}
  #bs-tabla{width:100%%;border-collapse:collapse;}
  #bs-tabla th,#bs-tabla td{text-align:left;padding:.45rem .6rem;border-bottom:1px solid #eee;vertical-align:top;}
  #bs-tabla thead th{border-bottom:2px solid #1E4A38;color:#1E4A38;}
  #bs-tabla tr:hover td{background:#f4f8f5;}
  #bs-conteo{color:#666;font-size:.85rem;margin:.3rem 0 1rem;}
</style>
<div class="bs-stats">
  <div class="bs-stat"><span class="n">%d</span><span class="l">accessions</span></div>
  <div class="bs-stat"><span class="n">%d</span><span class="l">families</span></div>
  <div class="bs-stat"><span class="n">%d</span><span class="l">species</span></div>
</div>
<div class="bs-controles">
  <div><label>Search</label><input type="text" id="bs-q" placeholder="name or species..."></div>
  <div><label>Family</label><select id="bs-fam">%s</select></div>
  <div><label>Origin</label><select id="bs-ori">%s</select></div>
  <div><label>Conservation</label><select id="bs-con">%s</select></div>
</div>
<div id="bs-conteo"></div>
<div class="tabla-scroll">
<table id="bs-tabla"><thead><tr><th>Id</th><th>Family</th><th>Scientific name</th><th>Common name</th><th>Origin</th><th>Conservation</th></tr></thead><tbody>',
  n_acc, n_fam, n_esp,
  opciones(sort(unique(bs$familia[bs$familia != "—"]))),
  opciones(sort(unique(bs$ori[nzchar(bs$ori)]))),
  sel_con))

for (i in seq_len(nrow(bs))) {
  r <- bs[i, ]
  buscar <- tolower(iconv(paste(r$nombre_comun, r$nombre_cientifico, r$id), to = "ASCII//TRANSLIT"))
  sci <- sci_html(r$nombre_cientifico)
  sci_cell <- if (nzchar(r$slug)) sprintf('<a href="fichas/%s.html">%s</a>', esc(r$slug), sci) else sci
  cat(sprintf('<tr data-fam="%s" data-ori="%s" data-con="%s" data-buscar="%s"><td>%s</td><td>%s</td><td>%s</td><td>%s</td><td>%s</td><td>%s</td></tr>',
    esc(r$familia), esc(r$ori), esc(r$iucn), esc(buscar),
    esc(r$id), esc(r$familia), sci_cell, esc(r$nombre_comun), esc(r$ori), r$con_html))
}

cat('</tbody></table>
</div>
<script>
(function(){
  var q=document.getElementById("bs-q"), ff=document.getElementById("bs-fam"),
      fo=document.getElementById("bs-ori"), fc=document.getElementById("bs-con"),
      filas=document.querySelectorAll("#bs-tabla tbody tr"), conteo=document.getElementById("bs-conteo");
  function aplica(){
    var qs=(q.value||"").toLowerCase(), vf=ff.value, vo=fo.value, vc=fc.value, n=0;
    filas.forEach(function(tr){
      var ok=tr.dataset.buscar.indexOf(qs)>=0 && (!vf || tr.dataset.fam===vf) && (!vo || tr.dataset.ori===vo) && (!vc || tr.dataset.con===vc);
      tr.style.display=ok?"":"none"; if(ok)n++;
    });
    conteo.textContent=n+" of "+filas.length+" accessions";
  }
  [q,ff,fo,fc].forEach(function(e){e.addEventListener("input",aplica);e.addEventListener("change",aplica);});
  aplica();
})();
</script>')
```

## Requesting or exchanging seeds

As part of the botanic-garden network (**ArbNet** accreditation and **BGCI** registration),
we make our collection available to institutions, researchers and restoration projects
through an **Index Seminum** (a list of seeds available for exchange). To request material,
propose an exchange or donate seeds of regional species, get in touch via our
[contact](contact.qmd) page.

---

*Source: list of seeds and cuttings from Estancia La Constancia. Conservation status from the
[IUCN Red List](https://www.iucnredlist.org). Updated by editing `data/banco_semillas.xlsx`.*
